Parsing Postfix log files

### 2009-02-27 doc/logparser-architecture.tex
Divided rules section in architecture into subsections.

XXX REWRITE\@: MERGE CONTENT FROM ELSEWHERE IN THE THESIS AND PAPER\@.

@@

XXX TO BE WRITTEN --- CHECK THE IMPLEMENTATION SECTION\@.

Each action is a separate procedure written to process a particular
category of input, e.g.\ rejections.  There may be many input variants
within one input category; in general each action will handle one category
of inputs, with each rule rule recognising one input variant.  It is
anticipated that parsers based on this architecture will have a high ratio
of rules to actions, with the aim of having simple rules and clear
distinctions between the inputs parsed by different rules.  An action may
process different input variants in slightly different ways, but large
changes in processing indicate the need for a new action and a new category
of input; if an action becomes unnecessarily complicated is starts to turn
into a monolithic parser, with too much logic contained in a single
procedure.

@@

During development of the Postfix log parser it became apparent that in
addition to the obvious variety in log \empty{fool check}lines there were
many complications to be overcome.  Some were the result of deficiencies in
Postfix's logging (some of which were rectified by later versions of
Postfix, e.g.\ identifying bounce
notifications~\sectionref{identifying-bounce-notifications}); others were
due to the vagaries of process scheduling, client behaviour, and
administrative actions.  All were successfully accommodated in the Postfix
log parser: adding new actions was enough to overcome several of the
complications; others required modifications to a single existing action to
work around the difficulties; the remainder were resolved by adapting
existing actions to cooperate and exchange extra data (via the framework),
changing their behaviour as appropriate based on that extra data.  Every
architecture should aim to make the hard things possible as well as making
the easy things easy (XXX MAYBE QUOTE LARRY\@?); the successful
implementation of the Postfix log parser demonstrates that this
architecture achieves that aim.

@@

@@

XXX TO BE WRITTEN\@.  TAKE CONTENT FROM THE PAPER AND POSSIBLY FROM THE
IMPLEMENTATION SECTION\@.

XXX DO I NEED TO SUBDIVIDE THIS INTO SUBSECTIONS\@?

@@

@@

Decoupling the rules from the actions and framework enables other rule
management approaches to be used, e.g.\ instead of manually adding new
rules, machine learning techniques could be used to automatically generate
new rules.  If this approach was taken the choice of machine learning
technique would be constrained by the size of typical data sets (see
\sectionref{parser efficiency}).  Techniques requiring the full data set
when training would be impractical; Instance Based
Learning~\cite{instance-based-learning} techniques that automatically
determine which inputs from the training set are valuable and which inputs
can be discarded might reduce the data required to a manageable size.  A
parser might also dynamically create new rules in response to certain
inputs, e.g.\ diagnostic messages indicating that the program which
produced the input being parsed had read a new configuration file, and had
changed its behaviour appropriately.  These avenues of research and
development have not been pursued by the author, but but the architecture
allows them to easily be undertaken independently.

@@

@@

@@

@@

Rule have certain characteristics that may help in understanding the
parser architecture:

XXX IMPROVE THIS\@; SHOULD IT BE A SUBSECTION\@?  EXPAND ON ALL THESE, OR
MOVE THEM TO FIT IN ELSEWHERE\@.  ADD MORE IF POSSIBLE\@.

\begin{itemize}

    \item The first matching rule wins: no further rules are tried against
        that input, but there is a mechanism for prioritising the rules so
        that more specific rules can be tried first.  XXX MOVE THIS TO THE
        PRIORITY SECTION\@.

    \item Rules are completely self-contained and can be understood in
        isolation, without reference to any other rules.  This is similar
        to a \acronym{CFG}; in \acronym{CFG} terms the parser rules could
        be described as:

@@

\end{itemize}

>>>>

@@

Each action is a separate procedure written to process a particular
category of input, e.g.\ rejections.  There may be many input variants
within one input category; in general each action will handle one category
of inputs, with each rule recognising one input variant.  It is anticipated
that parsers based on this architecture will have a high ratio of rules to
actions, with the aim of having simple rules and clear distinctions between
the inputs parsed by different rules.  An action may process different
input variants in slightly different ways, but large changes in processing
indicate the need for a new action and a new category of input; if an
action becomes unnecessarily complicated is starts to turn into a
monolithic parser, with too much logic contained in a single procedure.

@@

During development of the Postfix log parser it became apparent that in
addition to the obvious variety in log \empty{fool check}lines there were
many complications to be overcome.  Some were the result of deficiencies in
Postfix's logging (some of which were rectified by later versions of
Postfix, e.g.\ identifying bounce
notifications~\sectionref{identifying-bounce-notifications}); others were
due to the vagaries of process scheduling, client behaviour, and
administrative actions.  All were successfully accommodated in the Postfix
log parser: adding new actions was enough to overcome several of the
complications; others required modifications to a single existing action to
work around the difficulties; the remainder were resolved by adapting
existing actions to cooperate and exchange extra data (via the framework),
changing their behaviour as appropriate based on that extra data.  Every
architecture should aim to make the hard things possible as well as making
the easy things easy; the successful implementation of the Postfix log
parser demonstrates that this architecture achieves that aim.

@@

Glucose: $C_{6}H_{12}O_{6}$
Hydrocarbon: $CH_{3}{(CH_{2})}_{50}CH_{3}$

@@

@@

\subsection{Adding New Rules}

@@

Decoupling the rules from the actions and framework enables other rule
management approaches to be used, e.g.\ instead of manually adding new
rules, machine learning techniques could be used to automatically generate
new rules.  If this approach was taken the choice of machine learning
technique would be constrained by the size of typical data sets (see
\sectionref{parser efficiency}).  Techniques requiring the full data set
when training would be impractical; Instance Based
Learning~\cite{instance-based-learning} techniques that automatically
determine which inputs from the training set are valuable and which inputs
can be discarded might reduce the data required to a manageable size.  A
parser might also dynamically create new rules in response to certain
inputs, e.g.\ diagnostic messages indicating that the program which
produced the input being parsed had read a new configuration file, and had
changed its behaviour appropriately.  These avenues of research and
development have not been pursued by the author, but the architecture
allows them to easily be undertaken independently.

\subsection{Attaching Conditions To Rules}

@@

\subsection{Prioritisation and Rule Ordering}

@@

The first matching rule wins: no further rules are tried against that
input, but there is a mechanism for prioritising the rules so that more
specific rules can be tried first.

\subsection{Overlapping Rules}

@@

\subsection{Pathological Rules}

@@

    Rules are completely self-contained and can be understood in
        isolation, without reference to any other rules.  This is similar
        to a \acronym{CFG}; in \acronym{CFG} terms the parser rules could
        be described as:

@@

### 2009-02-27 doc/logparser.tex
Added delay, delays, and dsn to the results table; they're automatically     extracted and saved because filter_regex() fiddles the existing regex     snippets. Updated the thesis with new result columns, including the comparisons in the     state of the art review. Changed how the acronyms glossary is formatted by adding a new style that uses     eqlist. A couple of minor fixes.
Proofread postfix daemons appendix and made corrections/improvements.
Remove the space between glossary groups, so the three glossaries look better.
Proofread the glossary and made corrections. Removed some unnecessar

% Acronyms and glossary entries.  Useful options:
% * nonumberlist disables the list of page numbers after each entry.
\usepackage[acronym,toc,numberedsection=nolabel,style=list,nonumberlist]{glossaries}
\renewcommand{\glspostdescription}[0]{}
\renewcommand{\glsautoprefix}[1]{glossary:#1}
\newglossary[plg]{postfix}{pin}{pout}{Postfix Daemons}
\makeglossaries{}
\newcommand{\acronym}[1]{%
    \gls{#1}\glsadd{#1 glossary}%
}

% Provides commands to distinguish between pdf and dvi output.
\usepackage{ifpdf}
% When creating a \PDF{} make the table of contents into links to the pages
% (without horrible red borders) and include bookmarks.  The title and
% author do not work - I think either gnuplot or graphviz clobbers it.
% hyperfootnotes need to be disabled to avoid breaking footmisc, but they
% still seem to work, somehow.
\ifpdf{}
    \usepackage[pdftex,hyperfootnotes=false,plainpages=false,pdfpagelabels]{hyperref}
\else{}
    \usepackage[dvips,hyperfootnotes=false,plainpages=false,pdfpagelabels]{hyperref}
    % This is necessary for wrapping URLs in the bibliography when
    % producing a dvi, but causes problems when generating \PDF{} output.
    \usepackage{breakurl}
\fi{}
\hypersetup{
    pdftitle    = {Parsing Postfix log files},
    pdfauthor   = {John Tobin},
    final       = true,
    pdfborder   = {0 0 0},
}

@@

% Add last checked dates to all URLs
\newcommand{\urlLastChecked}[3][ ]{%
    \url{#2}#1(last checked #3)%
}

@@

\newcommand{\numberOFruleINTERSECTIONS}[0]{%
    14196%
}

@@

XXX EXTRACT DELAY AND DEFER WHEN AUTOMATIC EXTRACTION HAS BEEN FINISHED,
THEN UPDATE FIELDS IN BOTH DATABASE AND REVIEW OF OTHER PARSERS\@.

@@

\renewcommand{\glossarypostamble}{\label{Glossary}}
\printglossary{}
\renewcommand{\glossarypostamble}{\label{Acronyms}}
\printglossary[type=\acronymtype]{}
\renewcommand{\glossarypostamble}{\label{Postfix Daemons}}
\printglossary[type=postfix]{}

>>>>

% Acronyms and glossary entries.  Useful options:
% * nonumberlist disables the list of page numbers after each entry.
\usepackage[acronym,toc,numberedsection=nolabel,style=list,nonumberlist]{glossaries}
\renewcommand{\glspostdescription}[0]{}
\renewcommand{\glsautoprefix}[1]{glossary:#1}
\newglossary[plg]{postfix}{pin}{pout}{Postfix Daemons}
\makeglossaries{}
\newcommand{\acronyms}[1]{%
    \acronym[s]{#1}%
}
\newcommand{\acronym}[2][]{%
    \gls{#2}#1%
    % If there's a glossary entry for this acronym add it, otherwise
    % do nothing.
    \ifglsentryexists{#2 glossary}%
        {\glsadd{#2 glossary}}%
        {}%
}
% Define a new glossary style that uses eqlist, so that acronyms line up
% nicely.
\newglossarystyle{eqlist}{%
    \glossarystyle{list}%
    \renewenvironment{theglossary}{\begin{eqlist}}{\end{eqlist}}%
    \renewcommand*{\glossaryentryfield}[5]{%
        \item [\textbf{##2}] ##3%
    }%
    \renewcommand{\glsgroupskip}[0]{}%
}
% Define a new glossary style that removes the space between glossary
% groups.
\newglossarystyle{nospacelist}{%
    \glossarystyle{list}%
%    \renewenvironment{theglossary}{\begin{eqlist}}{\end{eqlist}}%
%    \renewcommand*{\glossaryentryfield}[5]{%
%        \item [\textbf{##2}] ##3%
%    }%
    \renewcommand{\glsgroupskip}[0]{}%
}

% Provides commands to distinguish between pdf and dvi output.
\usepackage{ifpdf}
% When creating a \PDF{} make the table of contents into links to the pages
% (without horrible red borders) and include bookmarks.  The title and
% author do not work - I think either gnuplot or graphviz clobbers it.
% hyperfootnotes need to be disabled to avoid breaking footmisc, but they
% still seem to work, somehow.
\ifpdf{}
    \usepackage[pdftex,hyperfootnotes=false,plainpages=false,pdfpagelabels]{hyperref}
\else{}
    \usepackage[dvips,hyperfootnotes=false,plainpages=false,pdfpagelabels]{hyperref}
    % This is necessary for wrapping URLs in the bibliography when
    % producing a dvi, but causes problems when generating \PDF{} output.
    \usepackage{breakurl}
\fi{}
\hypersetup{
    pdftitle    = {Parsing Postfix log files},
    pdfauthor   = {John Tobin},
    final       = true,
    pdfborder   = {0, 0, 0},
}

@@

% Add last checked dates to all URLs
\newcommand{\urlLastChecked}[3][ ]{%
    \url{#2}#1(last checked #3)%
}
\newcommand{\daemonDocURL}[3][]{%
    \hfill{} \newline{} \url{#2}%
    \hfill{} #1 Last checked #3%
}
\newcommand{\urlLastCheckedNoParens}[3][ ]{%
    \url{#2}#1last checked #3%
}

@@

\newcommand{\numberOFruleINTERSECTIONS}[0]{%
    17020%
}

@@

@@

\renewcommand{\glossarypostamble}{\label{Glossary}}
\printglossary[style=nospacelist]{}
\renewcommand{\glossarypostamble}{\label{Acronyms}}
\printglossary[type=\acronymtype,style=eqlist]{}
\renewcommand{\glossarypostamble}{\label{Postfix Daemons}}
\printglossary[type=postfix,style=nospacelist]{}
